\section{Recommendations for League Organizers}\label{sec:recommendations}

The purpose of this section is to aid league commissioners who wish to use the information presented in this paper to improve parity in their recreational leagues. Our work reveals that a player's experience in club leagues, where there is a vetting process to determine if a player can participate at their preferred division of competition, is more predictive than their own self-assessment as well as the assessments of any captains they have had in the past. For this reason, our first recommendation is that leagues use club ratings when they are available. 

Captain-assessment ratings should be used when club ratings are not present. This is a key result for regions that do not have the participation numbers to run club leagues like BUDA. For smaller organizations that are frustrated by the frequent occurrence of imbalanced teams, we highly recommend that captains regularly rate the players on their recreational league teams. Captain-assessment ratings are significantly more predictive of true skill level and future team performance than self-assessment ratings. League commissioners might even consider providing a financial reward to captains that complete player assessments at the end of each season, perhaps a refund of the registration fee or discount for future league participation. 

It is inevitable that some fraction of the ultimate frisbee population in a given league will not have a club-assessment rating or a captain-assessment rating. For these players, we recommend the use of self-assessment ratings.

It is important to emphasize that maximizing the value of club-assessment ratings requires that leagues continuously track and store roster data and game scores. Furthermore, club-assessment ratings will have the most utility if there are stratified divisions within the club league, such that a player who participates in one division is clearly better or worse (on average) than a player who participates in another division. The appropriate base rating for each division should be tailored to each league. We recommend using the Elo system as we have done here, and then translating the resulting Elo rating into the rating system in use by the league in question. The translation process should also be tailored to each league. To aid league organizers, the code we have used to build the club-assessment ratings for BUDA is publicly available \cite{shanegit}.