\section{Introduction}

Almost 5 million people in the Unitied States participate in the sport of Ultimate Frisbee.  Of these, 1.3 million are core participants---meaning that they play at least 13 times each year.  This is comparable to the sports of lacrosse (0.9 million) and rugby (0.4 million) combined \cite{sfia_2016}. Moreover, USA Ultimate---the premier organization for ultimate frisbee in the United States---has seen a 60\% growth in youth participation over the past five years \cite{usau_2016}.  While this segment of the population is focused on highly competitive play, it should still serve as a positive indicator for future growth of the sport.

Ultimate's emphasis on fair play is codified directly in the rulebook in the form of ``Spirit of the Game``.  From Section 1. Introduction, item B:

\begin{quote} Spirit of the Game. Ultimate relies upon a spirit of sportsmanship that places the responsibility for fair play on the player. Highly competitive play is encouraged, but never at the expense of mutual respect among competitors, adherence to the agreed upon rules, or the basic joy of play. Protection of these vital elements serves to eliminate unsportsmanlike conduct from the Ultimate field. Such actions as taunting opposing players, dangerous aggression, belligerent intimidation, intentional infractions, or other ``win-at-all-costs'' behavior are contrary to the Spirit of the Game and must be avoided by all players.
\end{quote}

What is a hat league?  How many hat leagues are there? Over 200.

What is a club league?  How many club leagues are there? About 20?

What is BUDA?

Why spring season?

Why the past 7 years?